---
title: "Analyzing Ransomware Data in R"
author: "Chris Campbell"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: '3'
    number_sections: true
  pdf_document:
    toc: true
    toc_depth: '3'
  word_document:
    toc: true
    toc_depth: '3'
always_allow_html: true
---

# Introduction

This project delves into analyzing ransomware infections using data extracted from the Shodan API. By analyzing real-time data on internet-connected devices, we explore ransomware trends across various countries and cities. Through data visualizations and statistical analysis, we aim to identify geographic hotspots of ransomware activity, comprehend infection patterns, and provide valuable insights for cybersecurity professionals. The project underscores the importance of monitoring and comprehending ransomware incidents to enhance global cyber defenses.


```{r libraries, setup, message=F, echo=T, results='hide', include=FALSE}
## Dependencies

# -   **R version**: `r R.version.string`
# -   **Required Libraries**: `dplyr`, `ggplot2`, `tidyr`, `httr2`, `stringr`,
# `ggthemes`, `renv`, `plotly`, `htmltools`, `maps`, `knitr`, `kableExtra`,
# `jsonlite`, `lintr`, `glue`, `virdis`


# Load required libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(httr2)
library(stringr)
library(ggthemes)
library(renv)
library(jsonlite)
library(plotly)
library(htmltools)
library(maps)
library(knitr)
library(kableExtra)
library(lintr)
library(glue)
library(viridis)
```

## Shodan API Overview

The Shodan API, a powerful tool for searching and retrieving data on internet-connected devices, provides information about devices’ locations, services, vulnerabilities, and more. In this project, the API is used to analyze global trends and patterns of ransomware infections.

-   [Shodan API Documentation ](https://developer.shodan.io/api)

```{r api-query, include=FALSE}
## Query Data from Shodan API

# Shodan api key and endpoint
api_key <- Sys.getenv("SHODAN_API_KEY") # Enter your API key here
api_url <- "https://api.shodan.io/shodan/host/search"

# Parameters to query
params <- list(
  key = api_key,
  query = "has_screenshot:true encrypted", # ransomware related query
  limit = 200 # Limit the number of results
)
```


```{r api-request, include=FALSE}
## Retrieve Data from Shodan API

# Send the request with httr2
request <- request(api_url) |>
  req_url_query(!!!params) |> # !!! Splices into key value pairs
  req_throttle(1) # API rate limit

# Perform a request to get the response
response <- req_perform(request)

# Parse the JSON
data <- resp_body_json(response, check_type = T, simplifyVector = T)

# Convert into a dataframe
shodan_df <- data.frame(data$matches)
```


```{r tidy-data, message=FALSE, include=FALSE, results='hide'}
# Tidy Data

# Select interesting columns
shodan_df_ransomware <- shodan_df %>%
  select(ip_str, port, transport, product, os, location, screenshot)
# Unnest nested columns
shodan_df_ransomware <- shodan_df_ransomware %>%
  unnest(`screenshot`) %>%
  unnest(`location`)

# Show Column names
colnames(shodan_df_ransomware)

# Select interesting columns from unnested dataframe
shodan_df_ransomware <- shodan_df_ransomware %>%
  select(
    ip_str, port, transport, product, os, country_name, country_code, city,
    longitude, latitude, text
  )

# Rename the columns
colnames(shodan_df_ransomware) <- c(
  "IP Address", "Port", "Transport", "Service",
  "Operating System", "Country", "Country Code",
  "City", "Longitude", "Latitude", "Ransom Letter"
)

# Group by Country Code
shodan_df_ransomware <- shodan_df_ransomware %>%
  filter(!is.na(`Ransom Letter`)) %>% # Only keep rows with ransom letters
  group_by(`Country Code`) %>%
  # Arrange by Country
  arrange(Country)

# Write the data to a CSV file
write.csv(shodan_df_ransomware, "data/shodan_ransomware.csv", row.names = FALSE)
```

# Data Analysis of Ransomware Infections

This section analyzes ransomware infections. It starts with a summary of affected countries and reported incidents. A statistical analysis presents key metrics on infection distribution. The section concludes with a table detailing ransomware incidents by country and city, revealing geographic trends and high-infection areas.

```{r data-analysis, echo=FALSE, results='asis'}
# Frequency table of infections by country
common_country_count <- table(shodan_df_ransomware$Country)
common_country_count <- sort(common_country_count, decreasing = TRUE) # Sort the count in descending order
# Get the names of the counts
common_country_names <- names(common_country_count)
# Get the most common country
most_common_country <- common_country_names[common_country_count == max(common_country_count)]
# Unique Countries
num_countries <- length(unique(shodan_df_ransomware$Country))


# Frequency table of infections by country and city
common_country_city_count <- table(shodan_df_ransomware$Country, shodan_df_ransomware$City)
# Counts of just the cities
common_country_city_count <- as.data.frame(common_country_city_count)
# Removes frequencies that are zero
common_country_city_count <- common_country_city_count[common_country_city_count$Freq != 0, ]
# Sorts the dataframe by the first column
common_country_city_count <- common_country_city_count[order(-common_country_city_count$Freq), ]
# Rename the columns
colnames(common_country_city_count) <- c("Country", "City", "Number of Infections")


# Most infections per city
common_city_count <- table(shodan_df_ransomware$City)
common_city_count <- sort(common_city_count, decreasing = TRUE)
# Names of the count
common_city_names <- names(common_city_count)
# The most common city
most_common_city <- common_city_names[common_city_count == max(common_city_count)]



# Initialize summary variables globally
country_text <- ""
city_text <- ""

# Construct the summary text
summary_text <- if (length(most_common_country) > 1 | length(most_common_city) > 1) {
  # If there are multiple countries or cities
  country_text <- paste(most_common_country, collapse = ", ")
  city_text <- paste(most_common_city, collapse = ", ")
  glue("
## Ransomware Infections Summary

According to the Shodan dataset, a total of **{nrow(shodan_df_ransomware)}** ransomware infections have been reported worldwide, impacting **{num_countries}** countries. The countries with the highest number of ransomware infections are **{country_text}**, each reporting **{max(common_country_count)}** incidents.

The cities with the most ransomware infections are **{city_text}**, each with **{max(common_city_count)}** incidents.

### Statistical Analysis
- The **average** number of ransomware infections per country is **{round(mean(common_country_count), 2)}**.
- The **median** number of ransomware infections per country is **{median(common_country_count)}**.
- The **standard deviation** of ransomware infections per country is **{round(sd(common_country_count), 2)}**.
")
} else {
  # If there is only one most common country or city
  country_text <- most_common_country
  city_text <- most_common_city
  glue("
## Ransomware Infections Summary

According to the Shodan dataset, a total of **{nrow(shodan_df_ransomware)}** ransomware infections have been reported worldwide, impacting **{num_countries}** countries. **{country_text}** has the highest number of ransomware infections, reporting **{max(common_country_count)}** incidents.

The city with the most ransomware infections is **{city_text}**, with **{max(common_city_count)}** incidents.

### Statistical Analysis
- The **average** number of ransomware infections per country is **{round(mean(common_country_count), 2)}**.
- The **median** number of ransomware infections per country is **{median(common_country_count)}**.
- The **standard deviation** of ransomware infections per country is **{round(sd(common_country_count), 2)}**.
")
}

# Display the summary
cat(summary_text, "\n")
```

## Table of Ransomware Infections by Country and City

This comprehensive table offers a detailed breakdown of ransomware infection rates across various countries and cities. It presents country and city names alongside the corresponding number of ransomware incidents, making it easy to compare regions. This table serves as a crucial reference point for understanding global ransomware trends and identifying areas where cyber defenses may need reinforcement.

```{r data-table, echo=FALSE}
# Table of Ransomware Infections by Country
kable(common_country_city_count,
  caption = "Distribution of Ransomware Infections by Country and City",
  col.names = c("Country", "City", "Number of Infections"), format = "html"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(width = "100%", height = "500px") # Add a scroll box
```

# Data Visualization of Ransomware Infections

This section visualizes ransomware infection patterns globally. It maps incidents at country and city levels using Shodan API data, highlighting affected regions and trends. An interactive map lets users zoom in and examine infection details, making it useful for cybersecurity professionals and researchers.

```{r map-plot, eval=FALSE, fig.height=6, fig.width=10, include=FALSE, out.width="100%"}
# Count the number of times values in Country appear
shodan_count <- shodan_df_ransomware %>%
  group_by(`Country Code`, `City`, `Longitude`, `Latitude`) %>%
  count(Country)

# Create a world map of ransomware infections
m <- ggplot(shodan_count, aes(x = Longitude, y = Latitude, color = `Country`, size = n)) +
  borders("world", colour = "gray50", fill = "gray50") +
  # Remove Antarctica
  coord_cartesian(xlim = c(-180, 180), ylim = c(-60, 90)) +
  geom_point() +
  theme_map() +
  labs(
    title = "The Global Impact of Ransomware Infections",
    caption = "Source: Shodan API",
    x = "Longitude",
    y = "Latitude",
    color = "Country"
  ) +
  theme_fivethirtyeight() +
  # Remove the gridlines and axis labels
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none", # Removes the fill legend
    plot.title = element_text(hjust = 0.5)
  ) # Center the title
```

## Exploring Ransomware Hotspots

This data visualization explores the global distribution of ransomware infections, focusing on the geographical hotspots by country and city. Using data from the Shodan API, the map highlights areas with the highest concentrations of ransomware incidents, shedding light on trends and patterns in cyberattacks. By mapping ransomware infections based on real-time data, the visualization provides insights into which regions are most affected and allows for a better understanding of the geographic spread of these cyber threats. The interactive map enables users to zoom in on specific locations and view detailed information on the number of incidents, cities, and countries impacted, offering valuable insights for cybersecurity professionals and researchers.

```{r interactive-map-plot, echo=FALSE}
# Count the number of times values in Country appear
shodan_count <- shodan_df_ransomware %>%
  group_by(`Country Code`, `City`, `Longitude`, `Latitude`) %>%
  count(Country)

# Make the map interactive
p <- ggplot(shodan_count, aes(
  x = Longitude, y = Latitude,
  color = n, size = n,
  text = paste(
    " City:", `City`, "<br>", "Country:", `Country`,
    "<br>", "Number of Infections:", n
  )
)) +
  borders("world", colour = "gray50", fill = "gray50") +
  # Remove Antarctica
  coord_cartesian(xlim = c(-180, 180), ylim = c(-60, 90)) +
  geom_point() +
  scale_color_viridis(option = "D", discrete = FALSE) + # Viridis color gradient
  theme_map() +
  labs(
    title = "Ransomware Hotspots Across the Globe",
    x = "Longitude",
    y = "Latitude",
    color = "Country"
  ) +
  theme_fivethirtyeight() +
  # Remove the gridlines and axis labels
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none", # Removes the fill legend
    plot.title = element_text(hjust = 0.5)
  ) # Center the title

ggplotly(p, tooltip = "text", width = 900, height = 600) %>%
  layout(
    autosize = TRUE,
    hoverlabel = list(align = "left"),
    # Create a caption for plot
    annotations = list(
      text = "Source: Shodan API",
      x = 1, y = -0.05, # Moves caption to right edge and below plot
      align. = "right", # Right-align ext
      showarrow = FALSE,
      xref = "paper", yref = "paper", # Keep caption relative to plot
      font = list(size = 12, color = "gray50")
    )
  ) %>%
  config(responsive = TRUE)
```
