---
title: "Parsing Data from Shodan API in R"
author: "Chris Campbell"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: '3'
    number_sections: true
  pdf_document:
    toc: true
    toc_depth: '3'
  word_document:
    toc: true
    toc_depth: '3'
always_allow_html: true
---

# Introduction

This R Markdown file explores data regarding internet-connected devices with via the Shodan API.

## Dependencies

-   **R version**: `r R.version.string`
-   **Required Libraries**: `dplyr`, `ggplot2`, `tidyr`, `httr`, `jsonlite`, 
`stringr`, `ggthemes`, `renv`, `plotly`, `htmltools`, `maps`

```{r setup, message=F, echo=T, results='hide'}
#Load required libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(httr)
library(jsonlite)
library(stringr)
library(ggthemes)
library(renv)
library(plotly)
library(htmltools)
library(maps)
```

# Input

-   **Shodan API Base URL**:(https://api.shodan.io)

## Query Data from Shodan API

```{r}
# Shodan api key and endpoint
api_key <- Sys.getenv("SHODAN_API_KEY") # Enter your API key here
api_url <- "https://api.shodan.io/shodan/host/search"

# Parameters to query
params <- list(
  key = api_key,
  query = "has_screenshot:true encrypted" # ransomware related query
)

```

# Retrieve Data from Shodan API

```{r message=F, echo=T, results='hide'}
# Retrieve up to 20 pages of the Shodan dataset
page <- 1  # Start with the first page
max_pages <- 20  # Set the new page limit to 20

# Begin the loop to retrieve the data
repeat {
  params$page <- page  # Set the page number

  # Print the page number
  cat("Retrieving page", page, "of the Shodan dataset...\n")
  flush.console()  # Ensures immediate output

  # Send the GET request
  response <- GET(api_url, query = params)

  # Return the error message if the status code is not 200
  if (response$status_code != 200) {
    stop(content(response, "text", encoding = "UTF-8"))
  }

  # Parse the JSON response
  shodan_data <- fromJSON(content(response, "text", encoding = "UTF-8"))

  # Create the dataframe
  shodan_df <- as.data.frame(shodan_data$matches)
  
  # Append the data from the current page to the existing dataframe
  if (page == 1) {
    shodan_df_all <- shodan_df
  } else {
    shodan_df_all <- bind_rows(shodan_df_all, shodan_df)
  }

  Sys.sleep(1)  # Sleep for 1 second to avoid rate limiting by Shodan
  
  # Stop after retrieving 20 pages  
  if (page >= max_pages) {
    message("Reached the maximum of 20 pages. Stopping.")
    break
  }

  page <- page + 1  # Increment the page number
}

```

# Tidy Data

```{r}
# Select interesting columns
shodan_df_ransomware <- shodan_df_all %>%
  select(ip_str, port, transport, product, os, location, screenshot)
# Unnest nested columns
shodan_df_ransomware <- shodan_df_ransomware %>%
  unnest(`screenshot`) %>%
  unnest(`location`)

# Show Column names
colnames(shodan_df_ransomware)

# Select interesting columns from unnested dataframe
shodan_df_ransomware <- shodan_df_ransomware %>%
  select(ip_str, port, transport, product,os, country_name, country_code, city, 
         longitude, latitude, text)

# Rename the columns
colnames(shodan_df_ransomware) <- c("IP Address", "Port", "Transport", "Service", 
                                    "Operating System", "Country", "Country Code",
                                    "City", "Longitude", "Latitude","Ransom Letter")

# Group by Country Code
shodan_df_ransomware <- shodan_df_ransomware %>%
  group_by(`Country Code`) %>%
  # Arrange by Country
  arrange(Country)


```

# Data Analysis of Ransomware Infections

```{r}
# Create a frequency table with the counts
common_country_count <- table(shodan_df_ransomware$Country)
common_country_count <- sort(common_country_count, decreasing = TRUE) # Sort the count in descending order

# Count the number of times values in Country appear
shodan_df_ransomware_count <- shodan_df_ransomware %>%
  count(Country)

# Convert into a factor
shodan_df_ransomware_count$`Country Code` <- 
  factor(shodan_df_ransomware_count$`Country Code`,
         levels = shodan_df_ransomware_count$`Country Code`)

# Get the names of the counts
common_country_names <- names(common_country_count)

# Get the most common country
most_common_country <- common_country_names[common_country_count == max(common_country_count)]
# Collapse the most common country into a single string
most_common_country <- paste(most_common_country, collapse = ", ")


# Output the most common country
cat("Per the Shodan dataset,", most_common_country,
    "is the country with the most ransomware infections.", "\n",
    "There are", max(common_country_count),"infections currently.", "\n")

# Display the total number of ransomware infections
cat("The total number of ransomware infections:", nrow(shodan_df_ransomware), 
    "\n")

```

# Data Visualization of Ransomware Infections

## Ransomware Infections by Country and City (ggplot2)

```{r}
# Create a world map of ransomware infections
ggplot(shodan_df_ransomware, aes(x = Longitude, y = Latitude, color = `City`)) +
  borders("world", colour = "gray50", fill = "gray50") +
  # Remove Antarctica
  coord_quickmap(xlim = c(-180, 180), ylim = c(-60, 90)) +
  geom_point() +
  theme_map() +
  labs(title = "Ransomware Infections by Country and City",
       caption = "Source: Shodan API",
       x = "Longitude",
       y = "Latitude",
       color = "Country Code") +
  theme_fivethirtyeight() +
  # Remove the gridlines and axis labels
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none", # Removes the fill legend
        plot.title = element_text(hjust = 0.5)) # Center the title
```

## Interactive Ransomware Infections Map (plotly)

```{r}
# Make the map interactive
p <- ggplot(shodan_df_ransomware, aes(x = Longitude, y = Latitude, color = `City`)) +
  borders("world", colour = "gray50", fill = "gray50") +
  geom_point() +
  theme_map() +
  labs(title = "Ransomware Infections by Country and City",
       caption = "Source: Shodan API",
       x = "Longitude",
       y = "Latitude",
       color = "Country Code") +
  theme_fivethirtyeight() +
  # Remove the gridlines and axis labels
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none", # Removes the fill legend
        plot.title = element_text(hjust = 0.5)) # Center the title

ggplotly(p)
```
